---
title: "Assignment 2 - Mapping"
author: "Taketo Muroya"
output: 
  html_document:
    code_folding: hide
    highlight: textmate
    toc: yes
always_allow_html: yes
---

```{r Setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path="figures/", cache.path="cache/", 
               cache=FALSE,echo=TRUE,message=FALSE,warning=FALSE)

```

```{r}
# Setting up
library(leaflet)
library(RColorBrewer)
library(sp)
library(rgeos)
library(geosphere)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(rgdal)
library(ggmap)
library(lubridate)
```

## 1. Location of Severe Fires

```{r}
# load data
rawfire <- read.csv("building_fires.csv")

# pick up highest severity (=5th alarm)
fire1 <- subset(rawfire,HIGHEST_LEVEL_DESC=="55 - Fifth Alarm")
fire2 <- subset(rawfire,HIGHEST_LEVEL_DESC=="5 - 5th alarm")
fire <- rbind(fire1, fire2)
# Initially 75 seems to be highest but followed http://www.fdnewyork.com/aa.asp
# fire <- subset(rawfire,HIGHEST_LEVEL_DESC=="75 - All Hands Working")

# create popup
content <- paste("When:",fire$INCIDENT_DATE_TIME,"<br/>",
                 "Where:",fire$address,"<br/>",
                 "Action:",fire$ACTION_TAKEN1_DESC,"<br/>")

# create base map
m <- leaflet(fire) %>%
  addTiles() %>% setView(-73.9326, 40.8075, zoom = 12) %>%
  addProviderTiles("Stamen.TonerLite")

# plot fire incidents
m %>% addCircleMarkers(lng = ~lon, lat = ~lat, popup = content)
```

## 2. Layers and Clusters
### a) Color by Type of Property

```{r}
# labeling property
fire$property <- substr(fire$PROPERTY_USE_DESC, 1, 1)
fire$property <- gsub("0", "Other", fire$property)
fire$property <- gsub("1", "Assembly", fire$property)
fire$property <- gsub("2", "Education", fire$property)
fire$property <- gsub("3", "Health", fire$property)
fire$property <- gsub("4", "Residence", fire$property)
fire$property <- gsub("5", "Business", fire$property)
fire$property <- gsub("6", "Manufacturing", fire$property)
fire$property <- gsub("7", "Manufacturing", fire$property)
fire$property <- gsub("8", "Other", fire$property)
fire$property <- gsub("9", "Outside", fire$property)
fire$property <- gsub("N", "Other", fire$property)
fire$property <- gsub("U", "Other", fire$property)

# create color and popup
pal = colorFactor("Set1", domain=fire$property)
color_offsel1 = pal(fire$property)

content <- paste("When:",fire$INCIDENT_DATE_TIME,"<br/>",
                 "Where:",fire$address,"<br/>",
                 "Action:",fire$ACTION_TAKEN1_DESC,"<br/>",
                 "Property:",fire$property,"<br/>")

# plot fire incidents by property
m %>% 
  addCircleMarkers(lng = ~lon, lat = ~lat, color = color_offsel1,
                   popup = content) %>% 
  addLegend(position = 'topleft', values = ~fire$property, 
            pal = pal, title = "Affected Property")
```

### b) Cluster

```{r}
m %>% 
  addCircleMarkers(clusterOptions=markerClusterOptions(),
                   color=color_offsel1, popup=content, 
                   lng = ~lon, lat = ~lat) %>%
  addLegend(position = 'topleft', values = ~fire$property, 
            pal = pal, title = "Affected Property")
```

## 3. Fire Houses

```{r}
house <- read.csv("FDNY_Firehouse_Listing.csv")

m %>% 
  addCircleMarkers(radius=fire$UNITS_ONSCENE/3, group="Incidents",
                   color=color_offsel1, popup=content, 
                   lng = ~lon, lat = ~lat) %>% 
  addLegend(position = 'topleft', values = ~fire$property, 
            pal = pal, title = "Affected Property") %>%
  addMarkers(data=house, group = "Firehouses",
             popup = ~as.character(house$FacilityName), 
             lng = ~Longitude, lat = ~Latitude) %>%
  addLayersControl(overlayGroups=c("Incidents","Firehouses"))
```

## 4. Distance from Firehouse and Response Time
### a) Calculate Distance

```{r}
house <- subset(house, house$Latitude != "NA")

# calculate distance from fire house
a <- SpatialPoints(rawfire[,c(27,26)])
b <- SpatialPoints(house[,c(5,6)])
rawfire$nearest <- 1000*apply(gDistance(a, b, byid=TRUE), 2, min)

# calculate response time
rawfire$happen <- parse_date_time(rawfire$INCIDENT_DATE_TIME,
                                  c('%m/%d/%Y %I:%M:%S %p'))
rawfire$arrival <- parse_date_time(rawfire$ARRIVAL_DATE_TIME,
                                   c('%m/%d/%Y %I:%M:%S %p'))
rawfire$response <- as.numeric(rawfire$arrival - rawfire$happen)

# cleaning data and omitting outliers
cleanfire <- subset(rawfire, rawfire$response != "NA")
cleanfire <- subset(cleanfire, cleanfire$response != "-305")
cleanfire <- subset(cleanfire, cleanfire$nearest < 1000)
cleanfire <- subset(cleanfire, cleanfire$response < 2000)

# create the high severity group
sev1 <- subset(cleanfire,HIGHEST_LEVEL_DESC=="55 - Fifth Alarm")
sev2 <- subset(cleanfire,HIGHEST_LEVEL_DESC=="5 - 5th alarm")
sev <- rbind(sev1, sev2)

# create the low severity group
nsev <- subset(cleanfire,HIGHEST_LEVEL_DESC=="0 - Initial alarm")

# regression analysis
all <- lm(response ~ nearest, data = cleanfire)
high <- lm(response ~ nearest, data = sev)
low <- lm(response ~ nearest, data = nsev)

# visualize the relationships
ggplot(cleanfire, aes(x=nearest,y=response)) + 
  geom_point() + stat_smooth(method="lm") + theme_economist() + 
  xlab("Distance from Firehouse") + ylab("Response time") + 
  ggtitle("Distance from Firehouse and Response Time (All)") + 
  geom_text(aes(30,850,label=paste("Slope =",round(all$coef[2],1))))

ggplot(sev, aes(x=nearest,y=response)) + 
  geom_point() + stat_smooth(method="lm") + theme_economist() + 
  xlab("Distance from Firehouse") + ylab("Response time") + 
  ggtitle("Distance from Firehouse and Response Time \n (High Severity: 5th alarm)") + 
  geom_text(aes(15,380,label=paste("Slope =",round(high$coef[2],1))))

ggplot(nsev, aes(x=nearest,y=response)) + 
  geom_point() + stat_smooth(method="lm") + theme_economist() + 
  xlab("Distance from Firehouse") + ylab("Response time") + 
  ggtitle("Distance from Firehouse and Response Time \n (Low Severity: Initial alarm)") + 
  geom_text(aes(15,380,label=paste("Slope =",round(low$coef[2],1))))
```

#### **Comment:** We can see positive slopes among these graphs. The slopes of low severity and high severity would be steeper than the total one. This may be because the distance from firehouse would be more likely to affect the response time in the low and high severe situation.

### b) Map of Response Times

```{r, results='hide'}
# load shape file of borough map
nyc_b <- readOGR("borough/.","nybb")
nyc_b <- spTransform(nyc_b, CRS("+proj=longlat +datum=WGS84"))
nyc_b <- fortify(nyc_b)

# load nyc map in google map
nycmap <- get_map("New York City", source="stamen",
                  maptype="toner-lite")
g <- ggmap(nycmap)
```

```{r}
# labeling property
cleanfire$property <- substr(cleanfire$PROPERTY_USE_DESC, 1, 1)
cleanfire$property <- gsub("0", "Other", cleanfire$property)
cleanfire$property <- gsub("1", "Assembly", cleanfire$property)
cleanfire$property <- gsub("2", "Education", cleanfire$property)
cleanfire$property <- gsub("3", "Health", cleanfire$property)
cleanfire$property <- gsub("4", "Residence", cleanfire$property)
cleanfire$property <- gsub("5", "Business", cleanfire$property)
cleanfire$property <- gsub("6", "Manufacturing", cleanfire$property)
cleanfire$property <- gsub("7", "Manufacturing", cleanfire$property)
cleanfire$property <- gsub("8", "Other", cleanfire$property)
cleanfire$property <- gsub("9", "Outside", cleanfire$property)
cleanfire$property <- gsub("N", "Other", cleanfire$property)
cleanfire$property <- gsub("U", "Other", cleanfire$property)

# aggregate by borough & property
pf <- aggregate(cleanfire$response, 
                by=list(cleanfire$BOROUGH_DESC, cleanfire$property),
                FUN=mean)
colnames(pf) = c("borough", "property", "response")

# combine borough map and aggregated data
pf$borough <- gsub("2 - Bronx", "Bronx", pf$borough)
pf$borough <- gsub("3 - Staten Island", "Staten Island", pf$borough)
pf$borough <- gsub("4 - Brooklyn", "Brooklyn", pf$borough)
pf$borough <- gsub("5 - Queens", "Queens", pf$borough)
pf$borough <- gsub("1 - Manhattan", "Manhattan", pf$borough)

nyc_b$id <- gsub("0", "Bronx", nyc_b$id)
nyc_b$id <- gsub("1", "Staten Island", nyc_b$id)
nyc_b$id <- gsub("2", "Brooklyn", nyc_b$id)
nyc_b$id <- gsub("3", "Queens", nyc_b$id)
nyc_b$id <- gsub("4", "Manhattan", nyc_b$id)

nyc_p <- left_join(pf, nyc_b, by = c("borough" = "id"))

# visualization by property
g +
  geom_polygon(aes(x=long, y=lat, group=group, fill=response), 
               color='black', size=0.0125, data=nyc_p) + 
  scale_fill_gradientn(colours=c("blue","purple2","red")) + 
  facet_wrap(~ property, ncol = 4)
```

#### **Comment:** The response time tends to be longer for the health property than for other properties, especially in Queens, Bronx, and Manhattan.

```{r}
# labeling severity
cleanfire$s <- substr(cleanfire$HIGHEST_LEVEL_DESC, 1, 1)
cleanfire$s <- gsub("0", "(0) Initial alarm", cleanfire$s)
cleanfire$s <- gsub("1", "(1) First alarm", cleanfire$s)
cleanfire$s <- gsub("2", "(2) Second alarm", cleanfire$s)
cleanfire$s <- gsub("3", "(3) Third alarm", cleanfire$s)
cleanfire$s <- gsub("4", "(4) Fourth alarm", cleanfire$s)
cleanfire$s <- gsub("5", "(5) Fifth alarm", cleanfire$s)
cleanfire$s <- gsub("7", "All Hands Working", cleanfire$s)

# aggregate by borough & severity
sf <- aggregate(cleanfire$response, 
                by=list(cleanfire$BOROUGH_DESC, cleanfire$s),
                FUN=mean)
colnames(sf) = c("borough", "severity", "response")

# combine borough map and aggregated data
sf$borough <- gsub("2 - Bronx", "Bronx", sf$borough)
sf$borough <- gsub("3 - Staten Island", "Staten Island", sf$borough)
sf$borough <- gsub("4 - Brooklyn", "Brooklyn", sf$borough)
sf$borough <- gsub("5 - Queens", "Queens", sf$borough)
sf$borough <- gsub("1 - Manhattan", "Manhattan", sf$borough)

nyc_s <- left_join(sf, nyc_b, by = c("borough" = "id"))

# visualization by severity
g +
  geom_polygon(aes(x=long, y=lat, group=group, fill=response), 
               color='black', size=0.0125, data=nyc_s) + 
  scale_fill_gradientn(colours=c("blue","purple2","red")) + 
  facet_wrap(~ severity, ncol = 4)
```

#### **Comment:** When we campare the response time between (0) Initial alarm and (5) Fifth alarm, the response time tends to be longer for initial, fourth, and fifth alarm in some areas.

```{r}
# aggregate by borough & year
cleanfire$year <- substr(cleanfire$happen, 1, 4)
af <- aggregate(cleanfire$response, 
                by=list(cleanfire$BOROUGH_DESC, cleanfire$year),
                FUN=mean)
colnames(af) = c("borough", "year", "response")

# combine borough map and aggregated data
af$borough <- gsub("2 - Bronx", "Bronx", af$borough)
af$borough <- gsub("3 - Staten Island", "Staten Island", af$borough)
af$borough <- gsub("4 - Brooklyn", "Brooklyn", af$borough)
af$borough <- gsub("5 - Queens", "Queens", af$borough)
af$borough <- gsub("1 - Manhattan", "Manhattan", af$borough)

nyc_y <- left_join(af, nyc_b, by = c("borough" = "id"))

# visualization by year
g +
  geom_polygon(aes(x=long, y=lat, group=group, fill=response), 
               color='black', size=0.0125, data=nyc_y) + 
  scale_fill_gradientn(colours=c("blue","purple2","red")) + 
  facet_wrap(~ year, ncol = 3)
```

#### **Comment:** The response time seems to be longer in all boroughs as the time goes, especially in 2018.